Modelling results

### 2024-09-05 notebooks/02-modelling-results.qmd
Finalised correlations, accuracy analysis, finer groups

:::: {.content-visible when-format="html"}
::: {.callout-note collapse="true"}
# Packages and setup
```r
{{< include ../scripts/_setup.R >}}
```
:::
::::

@@

@@

@@

The quality checks of the models are displayed in [@suppfig-implicit-checks and @suppfig-explicit-checks]{.content-visible when-format="html"} [@suppfig-implicit-checks-2 and @suppfig-explicit-checks-2]{.content-hidden when-format="html"}.

:::: {.content-visible when-format="html"}
::: {.panel-tabset .column-page-inset}

### Implicit task

@@

ggsave(here("figures/checks-explicit.png"), dpi = 600)
```

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the explicit task.](../figures/checks-explicit.png){#suppfig-explicit-checks}

:::
::::

:::{.content-hidden when-format="html"}

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the implicit task.](../figures/checks-implicit.png){#suppfig-implicit-checks-2}

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the explicit task.](../figures/checks-explicit.png){#suppfig-explicit-checks-2}

:::

@@

:::{#supptbl-glmm-i-summary}
```{r}
#| label: summary-implicit
#| echo: false

@@

:::{#supptbl-glmm-e-summary}
```{r}
#| label: summary-explicit
#| echo: false

@@

:::{#supptbl-estimates-implicit}
```{r}
#| label: emmeans-implicit
#| echo: false

@@

:::{#supptbl-estimates-explicit}
```{r}
#| label: emmeans-explicit
#| echo: false

@@

The figures below are also displayed in the main article. [See the HTML version for the code and details.] {.content-hidden when-format="html"}

@@

# Correlational analyses

:::: {.content-visible when-format="html"}

&nbsp;
&nbsp;
&nbsp;

@@

cat("═════════════════════════════════════════════════════════════════════════")
report_system(session = sessionInfo())
cat("Packages used:")
report_packages(session = sessionInfo())
cat("═════════════════════════════════════════════════════════════════════════")
```
:::
::::

>>>>

::: {.content-visible when-format="html"}
::: {.callout-note collapse="true"}
# Packages and setup

``` r
{{< include ../scripts/_setup.R >}}
```
:::
:::

@@

## Group differences

### Table

::: {#supptbl-questionnaire-results}
```{r}
#| label: table-questionnaires
#| echo: false

group_effects <- 
  df_questionnaires |>
  pivot_longer(
    cols = c("VVIQ":"SUIS"),
    names_to = "questionnaire",
    values_to = "score"
  ) |> 
  group_by(questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ aphantasia, data = data)),
    lm_ranked = list(lm(rank(score) ~ aphantasia, data = data)),
    ranked_means = list(prepare_means(lm_ranked, questionnaire)),
    table = list(generate_model_summary(lm, lm_ranked, questionnaire))
  )

bind_rows(group_effects$table) |> display()
```
:::

### Plot

```{r}
#| label: questionnaire-plot-setup
#| echo: false

# dodge width for all the geoms
dw <- 1
# axis text size
txt <- 14
# legend text size
txt_legend <- 12

norm_means <- 
  df_q_norm |> 
  group_by(aphantasia) |>
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "score"
  ) |> 
  group_by(Questionnaire) |> 
  nest() |> 
  rowwise() |> 
  mutate(
    lm = list(lm(score ~ aphantasia, data = data)),
    means = list(estimate_means(lm, by = "aphantasia"))
  ) |> 
  unnest(means) |> 
  select(!c(data, lm)) |> 
  ungroup() |> 
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire,
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    ))
```

::: {#suppfig-questionnaire-plot}
```{r}
#| label: plot-questionnaires
#| fig-width: 10
#| fig-height: 5

df_q_norm |>

  # Long format first -----------------------------------------------------
  pivot_longer(
    cols = c(VVIQ:SUIS),
    names_to = "Questionnaire",
    values_to = "Score"
  ) |>
  mutate(
    Questionnaire = fct_relevel(
      Questionnaire, 
      c("VVIQ", "SUIS", "OSIQ_Object", "OSIQ_Spatial")
    )) |>

  # Plotting --------------------------------------------------------------
  ggplot(aes(
    x = Questionnaire, 
    y = Score, 
    fill = aphantasia, 
    color = aphantasia
    )) +

  # accentuation of the median line at 0.5
  geom_hline(
    yintercept = .5,
    linetype = 1,
    color = "grey30",
    alpha = .2
  ) +

  # Scores ---------------------------------------------------------------
  # violin shapes for the distributions and quantiles
  geom_violin(
    position = position_dodge(dw),
    alpha = .1,
    draw_quantiles = c(.25, .5, .75)
  ) +
  # points for means and lines for CIs
  geom_pointrange(
    data = norm_means,
    aes(
      x = Questionnaire,
      y = Mean,
      ymin = CI_low,
      ymax = CI_high,
      color = aphantasia
    ),
    size = .75,
    linewidth = 1,
    position = position_dodge(dw)
  ) +

  # Stars and lines --------------------------------------------------------
  # stars for significance
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 1, y = .95, size  = 10 
    ) +
  # dotted line that separates the different questionnaires
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 1.5)),
    linetype = 3
  ) +
  # stars 
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 2, y = .95, size  = 10 
    ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 2.5)),
    linetype = 3
  ) +
  # stars
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 3, y = .95, size  = 10 
    ) +
  # dotted line
  geom_vline(
    aes(xintercept = stage(Questionnaire, after_scale = 3.5)),
    linetype = 3
  ) +
  # stars
  annotate(
    geom  = "text", label = "***", color = "black",
    x = 4, y = .95, size  = 10 
    ) +

  # Aesthetics -------------------------------------------------------------
  scale_color_okabeito(name = "Group:  ", labels = c(" Control   ", " Aphantasic")) +
  scale_fill_okabeito(name = "Group:  ", labels  = c(" Control   ", " Aphantasic")) +
  scale_x_discrete(
    name = "",
    labels = c("VVIQ", "SUIS", "OSIQ-Object", "OSIQ-Spatial")
  ) +
  scale_y_continuous(
    name = "Standardized score",
    breaks = breaks_pretty(8)
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(),
    axis.text.x  = element_text(size = txt),
    axis.title.y = element_text(size = txt),
    legend.title = element_text(size = txt_legend),
    legend.text  = element_text(size = txt_legend),
    legend.position = "top"
    )

# ggsave(here("figures/questionnaires.png"), dpi = 600)
```
:::

## Congruence effects correlations

### Matrices

```{r}
#| label: tables-corr-questionnaires

corr_whole <- 
  df_questionnaires |> 
    select(aphantasia, "Implicit effect":"SUIS") |>
    correlation(method = "spearman", p_adjust = "fdr")

corr_aph <-
  df_questionnaires |>
    select(aphantasia, "Implicit effect":"SUIS") |>
    filter(aphantasia == "yes") |>
    correlation(method = "spearman", p_adjust = "fdr")

corr_control <-
  df_questionnaires |>
    select(aphantasia, "Implicit effect":"SUIS") |>
    filter(aphantasia == "no")  |>
    correlation(method = "spearman", p_adjust = "fdr")

# corr_whole
# corr_aph
# corr_control
```

::: {#suppfig-corr-mat-questionnaires}
```{r}
#| label: corr-mat-questionnaires
#| fig-width: 10
#| fig-height: 6
#| eval: false

corr_whole |> summary() |> plot() + 
  labs(title = "Correlation matrix, whole group (Spearman FDR corrected)")

# ggsave(here("figures/corr-mat-whole.png"), dpi = 600)

corr_aph |> summary() |> plot() +
  labs(title = "Correlation matrix, aphantasia group only (Spearman FDR corrected)")

# ggsave(here("figures/corr-mat-aph.png"), dpi = 600)

corr_control |> summary() |> plot() +
  labs(title = "Correlation matrix, control group only (Spearman FDR corrected)")

# ggsave(here("figures/corr-mat-control.png"), dpi = 600)
```
:::

### Plots

```{r}
#| label: corr-plots-questionnaires-setup
#| fig-width: 10
#| fig-height: 5
#| include: false

note_size = 6

# Ranked versions
p_corr_vviq <-
  df_q_ranked |> 
  plot_correlations("VVIQ", "Implicit effect") +
  labs(x = "VVIQ (ranked)", y = "Implicit effect") +
  annotate(
    geom  = "text", 
    label = TeX(r"($\rho = 0.21, \textit{p} = 0.02$)"), 
    color = "black",
    x = 89, y = .25, 
    size  = note_size 
    )

p_corr_osiq <-
  df_q_ranked |> 
  plot_correlations("OSIQ_Object", "Implicit effect") +
  labs(x = "OSIQ-Object (ranked)", y = "Implicit effect") +
  annotate(
    geom  = "text", 
    label = TeX(r"($\rho = 0.26, \textit{p} = 0.003$)"), 
    color = "black",
    x = 80, y = .25, 
    size  = note_size 
    )

p_corr_suis <-
  df_q_ranked |> 
  plot_correlations("SUIS", "Implicit effect") +
  labs(x = "SUIS (ranked)", y = "Implicit effect") +
  annotate(
    geom  = "text", 
    label = TeX(r"($\rho = 0.27, \textit{p} = 0.002$)"), 
    color = "black",
    x = 81, y = .25, 
    size  = note_size 
    )
```

::: {#suppfig-corr-plots-questionnaires}
```{r}
#| label: corr-plots-questionnaires
#| fig-width: 20
#| fig-height: 15

layout <- "
AAAABBBB
##CCCC##
"

# ((p_corr_vviq + p_corr_osiq) / p_corr_suis) +
(p_corr_vviq + p_corr_osiq + p_corr_suis) +
  plot_layout(design = layout, guides = "collect") &
  theme(
    legend.position = "top", 
    legend.text = element_text(size = 20),
    legend.title = element_blank(),
    axis.title = element_text(size = 16)
    )

# ggsave(here("figures/corr-plots.png"), dpi = 600)
```
:::

@@

We will quickly check for accuracy differences between groups with a logistic regression model.

```{r}
#| label: accuracy-models

fit_acc_im <- 
  glmer(
    formula = correct_implicit ~ (aphantasia + congruence + color)^3 + (1 | subjectid),
    data = df_i_acc,
    family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa")
  )

fit_acc_ex <- 
  glmer(
    formula = correct_explicit ~ (aphantasia + congruence + color)^3 + (1 | subjectid),
    data = df_e_acc,
    family = binomial(link = "logit"),
    control = glmerControl(optimizer = "bobyqa")
  )
```

::: {#supptbl-acc-i-summary}
```{r}
#| label: summary-accuracy-implicit
#| echo: false

model_performance(fit_acc_im) |> display()

model_parameters(fit_acc_im) |> display()
```

Performance and estimates of the logistic GLMM fitted on accuracy in the implicit task.
:::

::: {#supptbl-acc-i-estimates}
```{r}
#| label: estimates-accuracy-implicit
#| echo: false

estimate_means(fit_acc_im, by = "aphantasia", transform = "response") |> display()

estimate_contrasts(fit_acc_im, contrast = "aphantasia", transform = "response", p_adjust = "none") |> display()
```

Estimated probability of a correct answer and log-odds contrasts between groups for the accuracy in the implicit task.
:::

::: {#supptbl-acc-e-summary}
```{r}
#| label: summary-accuracy-explicit
#| echo: false

model_performance(fit_acc_ex) |> display()

model_parameters(fit_acc_ex) |> display()
```

Performance and estimates of the logistic GLMM fitted on accuracy in the explicit task.
:::

::: {#supptbl-acc-e-estimates}
```{r}
#| label: estimates-accuracy-explicit
#| echo: false

estimate_means(fit_acc_ex, by = "aphantasia", transform = "response") |> display()

estimate_contrasts(fit_acc_ex, contrast = "aphantasia", transform = "response", p_adjust = "none") |> display()
```

Estimated probability of a correct answer and log-odds contrasts between groups for the accuracy in the explicit task.
:::

@@

The quality checks of the models are displayed in @suppfig-implicit-checks[ and @suppfig-explicit-checks]{.content-visible when-format="html"} @suppfig-implicit-checks-2[ and @suppfig-explicit-checks-2]{.content-hidden when-format="html"}.

::: {.content-visible when-format="html"}
::: {.panel-tabset .column-page-inset}
### Implicit task

@@

# ggsave(here("figures/checks-explicit.png"), dpi = 600)
```

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the explicit task.](../figures/checks-explicit.png){#suppfig-explicit-checks}
:::
:::

::: {.content-hidden when-format="html"}
![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the implicit task.](../figures/checks-implicit.png){#suppfig-implicit-checks-2}

![Model assumption checks for the Generalized Linear Mixed Model fit on the RTs in the explicit task.](../figures/checks-explicit.png){#suppfig-explicit-checks-2}
:::

@@

::: {#supptbl-glmm-i-summary}
```{r}
#| label: summary-implicit
#| echo: false

@@

::: {#supptbl-glmm-e-summary}
```{r}
#| label: summary-explicit
#| echo: false

@@

::: {#supptbl-estimates-implicit}
```{r}
#| label: emmeans-implicit
#| echo: false

@@

::: {#supptbl-estimates-explicit}
```{r}
#| label: emmeans-explicit
#| echo: false

@@

The figures below are also displayed in the main article. \[See the HTML version for the code and details.\] {.content-hidden when-format="html"}

@@

# Finer groups

```{r}
#| label: checking-finer-groups

df_questionnaires |> 
  group_by(aphantasia, sub_group) |> 
  count() |> 
  arrange(sub_group)
```

Interestingly, we have 39 participants that did not score at floor VVIQ but between 17 and 32, a score range that @reeder have proposed to call "hypophantasia". However, our sample comprised only 2 hyperphantasics, i.e. participants scoring above 74 [@zeman2024]. These two participants will therefore be removed from the sample before performing new analyses on the aphantasic, hypophantasic and typical groups.

```{r}

model_formula_finer <- rt ~ (sub_group + congruence + color)^3 + (1|subjectid)

fit_finer_i <- 
  glmer(
    formula = model_formula_finer,
    data = df_i_finer,
    family = inverse.gaussian(link = "identity"),
    control = glmerControl(optimizer = "bobyqa")
  )

fit_finer_e <-
  glmer(
    formula = model_formula_finer,
    data = df_e_finer,
    family = inverse.gaussian(link = "identity"),
    control = glmerControl(optimizer = "bobyqa")
  )
```

```{r}
#| label: checks
#| fig-width: 14
#| fig-height: 12
#| eval: false

check_model(
  fit_finer_i, 
  checks = model_checks, 
  detrend = FALSE, 
  residual_type = "normal"
  ) |> 
  plot()

# ggsave(here("figures/checks-finer-implicit.png"), dpi = 600)

check_model(
  fit_finer_e, 
  checks = model_checks, 
  detrend = FALSE, 
  residual_type = "normal"
  ) |> 
  plot()

# ggsave(here("figures/checks-finer-explicit.png"), dpi = 600)
```

```{r}
model_performance(fit_finer_i)

model_parameters(fit_finer_i)
```

```{r}
model_performance(fit_finer_e)

model_parameters(fit_finer_e)
```

```{r}
#| label: emmeans-finer
#| echo: false

Anova(fit_finer_i, type = 3)
Anova(fit_finer_e, type = 3)

# emmeans_im <-
  emmeans(fit_finer_i, c("congruence", "sub_group")) |>
  summary(point.est = median) |>
  as_tibble() |>
  mutate(across(3:7, ~.*1000)) |>
  rename(median = emmean)

# emmeans_im |> 
#   rename(Group = 1, Condition = 2, "Median (ms)" = median) |>
#   display()

fit_finer_i |> 
  estimate_contrasts(
    contrast = c("congruence", "sub_group"), 
    p_adjust = "none"
  )

# emmeans_ex <-
  emmeans(fit_finer_e, c("congruence", "sub_group")) |>
  summary(point.est = median) |>
  as_tibble() |>
  mutate(across(3:7, ~.*1000)) |>
  rename(median = emmean)

fit_finer_e |> 
  estimate_contrasts(
    contrast = c("congruence", "sub_group"), 
    p_adjust = "none"
  )
```

::: {.content-visible when-format="html"}

@@

cat("═════════════════════════════════════════════════════════════════════════")
report_system(session = sessionInfo())
cat("Packages used:")
report_packages(session = sessionInfo())
cat("═════════════════════════════════════════════════════════════════════════")
```
:::
:::
